Selective Sequential Model Selection

### 2015-12-08 paper.tex
fixed some minor things. This is arxiv v1

\newcommand{\cM}{\mathcal{M}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\FDR}{\textnormal{FDR}}
\newcommand{\FCR}{\textnormal{FCR}}
\newcommand{\crt}{\phi}
\newcommand{\M}{\mathcal{M}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\Gv}{\;\;\big|\;\;}
\newcommand{\proj}{\cP}
\newcommand{\pow}{\text{Pow}}
\newcommand{\supp}{\text{supp}}
\newcommand{\sF}{\mathscr{F}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\sC}{\mathscr{C}}
\newcommand{\hJ}{\widehat{J}}
\newcommand{\bH}{\mathbf{H}}
\newcommand{\bM}{\mathbf{M}}
\newcommand{\tM}{\widetilde{M}}
\newcommand{\tE}{\widetilde{E}}
\newcommand{\tV}{\widetilde{V}}
\newcommand{\tR}{\widetilde{R}}
\newcommand{\tL}{\widetilde{L}}
\newcommand{\hk}{\hat{k}}
\newcommand{\hr}{\hat{r}}       
\newcommand{\cN}{\mathcal{N}}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}
\newcommand{\Err}{\mathcal{E}}
\newcommand{\RSS}{\text{RSS}}

@@

Following \citet{fithian2014optimal}, we say that for a fixed candidate null model $M$, the random variable $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is stochastically larger than uniform (henceforth super-uniform) under sampling from any $F\in M$, given that $M$ is selected. That is,
\begin{equation*}
\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = M\right) 
\leq \alpha, \quad \forall F\in M, \; \alpha \in [0,1].
\end{equation*}
Once we have constructed selective $p$-values for each fixed $M$, we can use them as building blocks to construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
\[
p_k(y) = p_{k, M_{k-1}(y)}(y),
\]
which is super-uniform on the event $\{F \in M_{k-1}(Y)\}$:
\begin{equation}\label{eq:selectiveGuaranteePk}
\P_F\left(p_k \leq \alpha \mid F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
\end{equation}
One useful strategy for constructing exact selective tests is to condition on the sufficient statistics of the null model $M_{k-1}$. By sufficiency, the law
\[
\L_F(Y \;\mid\; U_{k-1}, \; M_{k-1}=M)
\]
is the same for every $F\in M$. Thus, we can construct selective tests and $p$-values by comparing the realized value of any test statistic $T_k(Y)$ to its known conditional distribution under the null. It remains only to choose the test statistic and compute its conditional null distribution, which can be challenging. See \citet{fithian2014optimal} for a general treatment. Section~\ref{sec:computation} discusses computational strategies for the tests we propose in this article.

@@

A natural choice of test statistic is the {\em greedy likelihood ratio statistic}
\begin{equation}\label{eq:greedyLRT}
G_k(Y) = \sup_{\theta\in \Theta(E_k)} {\ell(\theta; Y)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell(\theta; Y)},
\end{equation}
which is the generalized likelihood ratio statistic for testing $M(E_{k-1})$ against the ``greedy'' alternative with one more active parameter, $\bigcup_{j \notin E_{k-1}} M(E_{k-1} \cup \{j\})$. The {\em selective greedy likelihood ratio test} rejects for large $G_k$, based on the law
\begin{equation*}
\L\left(G_k \;\mid\; M_{k-1}, U_{k-1}\right)
\end{equation*}
Because the likelihood in linear regression is a monotone decreasing function of the residual sum of squares, the max-$t$ test is equivalent to the greedy likelihood ratio test. The counterpart of the max-$t$ test in linear regression with known $\sigma^2$ is the max-$z$ test, which differs only in replacing the $t$-statistics $t_{j,E}$ with their corresponding $z$-statistics. The max-$z$ test is also equivalent to the selective greedy likelihood ratio test.

@@

In this setting, $\lambda_k$ is a natural test statistic for model $M_{k-1}$, with larger values suggesting a poorer fit. The {\em selective next-entry test} is the test that rejects for large $\lambda_k$, based on the law
\begin{equation*}
\L\left(\lambda_k \;\mid\; U_{k-1}, M_{k-1}\right).
\end{equation*}

@@

We now review several proposals for ordered testing of fixed null hypotheses along with their error guarantees in the traditional setting, where the sequence of null hypotheses is fixed. In the next section we will extend the analysis to random null hypotheses.

@@

A proof of Proposition~\ref{prop:counterexample} is given in the appendix.  The problem is that $k_0$ is no longer a fixed index. Consequently, even though $\L(p_k \mid H_k \text{ true})=U[0,1]$ for each fixed $k$, the $p$-value \smash{$p_{k_0}$} corresponding to the {\em first} true null hypothesis is stochastically smaller than uniform. In this counterexample, $k_0=2 \iff |Y_3|>z_{\alpha/2}$, giving $\text{FWER}\eqAS 1$ on the event $k_0=2$ and leaving no room for error when $k_0\neq 2$.

@@

Similarly, the error control guarantees of \citet{gsell2013sequential} and \citet{li2015accumulation} do not directly apply to case where the null hypotheses are random. However, we recover these guarantees if we have {\em conditional} independence and uniformity of null $p$-values given $k_0$:
\begin{equation}\label{eq:indepCond_random_k0}
  \P_F(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
  \mid p_1, \ldots, p_k) \eqAS \left(\prod_{i=k+1}^d \alpha_i\right) \quad \text{on } \{k_0(Y, F) = k\}.
\end{equation}
For the sake of brevity, we will say that $p$-value sequences satisfying~\eqref{eq:indepCond_random_k0} are {\em independent on nulls}.

The following proposition shows that independence on nulls allows us to transfer the error-control guarantees of ForwardStop and accumulation tests to the random-hypothesis case.
\begin{proposition}
  Let $\hk$ be a stopping rule operating on $p$-values $p_{1:d}(Y)$ for nested hypotheses $H_{1:d}(Y)$. For some function $g$, let
  \[
  \Err_g = \E\left[ g\left(\hk(p_{1:d}(Y)), k_0(Y, F)\right)\right],
  \]
  Suppose that \smash{$\hk$} controls $\Err_g$ at level $\alpha$ if $H_{1:d}$ are fixed and the null $p$-values are uniform and independent of each other and the non-nulls. Then, \smash{$\hk$} controls $\Err_g$ at level $\alpha$ whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
\end{proposition}
\begin{proof}
    For nested hypotheses, $k_0$ completely determines the truth or falsehood of $H_k(Y)$ for every $k$. If \smash{$\hk$} controls $\Err_g$ in the 
fixed-hypothesis case, it must in particular control $\Err_g$ conditional on any fixed values of $p_{1:k_0}$. Thus, \eqref{eq:indepCond_random_k0} implies
  \[
  \E\left[ g\left(\hk(p_1,\ldots,p_d), k_0(Y, F)\right) \;\mid\;
      k_0(Y, F), \; p_{1:k_0}(Y)\right] \leqAS \alpha.
  \]
  Marginalizing over $(p_{1:k_0},k_0)$ gives $\Err_g\leq \alpha$.
\end{proof}

@@

Recall that at step $k$, the max-$t$ test rejects for large \smash{$T_k=|t_{E_{k-1}}^*|$}, comparing its distribution to the conditional law
\begin{equation}\label{eq:maxTnull}
\L\left(T_k \; \mid \; E_{k-1}, \,X_{E_{k-1}}'Y, \,\|Y\|^2 \right).
\end{equation}
This conditional distribution is the same for all $F\in M(E_{k-1})$, because \smash{$U_{k-1}=(X_{E_{k-1}}'Y,\; \|Y\|^2)$} is the complete sufficient statistic for $M(E_{k-1})$.

If $F\in M(E_{k-1})$, then $p_k$ is uniform and independent of the previous $p$-values $p_{1:(k-1)}$ since:
\begin{enumerate}
\item $p_k$ is uniform conditional on $E_{k-1}$ and $U_{k-1}$ by construction, and
\item $p_1, \ldots, p_{k-1}$ are functions of $E_{k-1}$ and $U_{k-1}$, as we will see shortly.
\end{enumerate}
Informally, the pair $(E_{k-1}, U_{k-1})$ forms a ``wall of separation'' between $p_k$ and $p_{1:(k-1)}$, guaranteeing that $\L(p_k \mid p_{1:(k-1)}) = U[0,1]$ whenever $H_k$ is true.

@@

Furthermore, for $i<k$, $p_i$ is computed by comparing $T_i=|t_{j_i, E_{i-1}}|$, which is a function of $(E_i, U_i)$, to the reference null distribution $\L_{H_i}(T_i \mid E_{i-1}, U_{i-1})$, which is a function of $(E_{i-1}, U_{i-1})$. Having verified that under $H_k$, $p_k$ is uniform and independent of $p_{1:{k-1}}$, we can apply this conclusion iteratively to see that all remaining $p$-values are also uniform and independent.

By contrast, the saturated model $p$-values use a law different from~\eqref{eq:maxTnull} to compute $p$-values, one that uses
information not contained in $(E_k,U_k)$. As a result, saturated-model $p$-values are generally not independent on nulls. We discuss the regression setting in more detail in Section \ref{sec:selective-reg}.

@@

\begin{corollary}\label{cor:greedyLikIndep}
  If $U(Y; M)$ is complete and sufficient for model $M$, then the selective greedy likelihood $p$-values computed for the forward stepwise path are independent on nulls.
\end{corollary}
\begin{proof}
  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the sufficient filtration for the forward stepwise path. First, the test statistic $G_k$ is $\sF_k$-measurable because
  \begin{align*}
    G_k(Y) &= \sup_{\theta\in \Theta(E_k)} {\ell(\theta; Y)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell(\theta; Y)}\\
    &= \sup_{\theta\in \Theta(E_k)} {\ell_{E_k}^{Y\mid U}(\theta; U_k)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell_{E_k}^{Y\mid U}(\theta; U_k)}.
  \end{align*}
  Second, the null distribution $\L\left(G_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $G_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
\end{proof}

@@

  Furthermore, for all $\lambda\in\Lambda, \lambda\geq \gamma_k$, we have \smash{$\tE_\lambda \sub E_k$}. As a result, \smash{${\hat\theta^{\lambda} =\hat\theta^{(E_k,\lambda)}}$} for such $\lambda$, so $\tE_{\lambda}$ depends only on \smash{$\ell_{E_k}^{Y\mid U}(Y;\, U_k)$}, and therefore $\lambda_k$ can be computed from $(E_k, U_k)$.  Second, the null distribution $\L\left(\lambda_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $\lambda_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
\end{proof}

@@

Let $\eta_k\in \R^n$ denote the vector for which \smash{$\theta_{j_k}^{E_k} = \eta_k'\mu$}, and let \smash{$\proj_{\eta_k}^\perp$} denote the projection operator into the space orthogonal to $\eta_k$. The UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution \smash{$\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)$}, or equivalently
\begin{equation}\label{eq:satModel}
\L\left( X_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)
\end{equation}

@@

\subsubsection{Testing Against the Next Model}\label{sec:identify}
One option for selected-model inference is to perform a selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$, conditioning on both models. Conditioning on $(M_{k-1},M_k)$ is equivalent to conditioning on $(E_{k-1}, j_k)$. If $\sigma^2$ is known, the test is based on
\begin{equation}\label{eq:selModel_cond}
\L\left(X_{j_k}'Y \mid E_{k-1}, \; j_k, \; X_{E_{k-1}}'Y\right).
\end{equation}
If $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$. We can construct $p$-values from the UMPU, equal-tailed, or one-sided test based on the one-parameter exponential family~\eqref{eq:selModel_cond}; the only difference is in how much of the overall level $\alpha$ is apportioned to the right or left tail \citep{fithian2014optimal}. If we use the construction in~\eqref{eq:selModel_cond} after a subpath-sufficient path algorithm, the resulting $p$-values are always guaranteed to have independence on nulls. This is because the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, as is the reference null distribution.

@@

To avoid conditioning on the next model $M_k$, we could instead test $M_{k-1}$ against the alternative $M_{\infty}\setminus M_{k-1}$. In that case, we could modify~\eqref{eq:selModel_cond} and base the test on
\begin{equation}\label{eq:selModel_marg}
\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right).
\end{equation}
Again, if $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$. We could also replace $X_{j_k}'Y$ with any other test statistic $T(Y)$. As long as $T(Y)$ is measurable with respect to $(M_k, U_k)$, and the path algorithm is subpath-sufficient, the $p$-values are guaranteed to be independent on nulls. The next-entry test for the lasso and the max-$t$ test in forward-stepwise regression with unknown $\sigma^2$ are both examples of this approach. Note that under resampling from the law in~\eqref{eq:selModel_marg}, the selected variable $j_k$ is random; by contrast, \eqref{eq:selModel_cond} conditions on $j_k$.

@@

\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\textwidth}
    % source code: bivariateSelVSat.R
    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_condSets.pdf}
    \caption{The selected-model test conditions on $j_1=1$ (yellow region), while the saturated-model test also conditions on $Y_2=2.5$ to eliminate the nuisance variable $\mu_2$ (brown line segments).}
    \label{fig:bv_condSets}
  \end{subfigure}
  \hspace{.1\textwidth}
  \begin{subfigure}[t]{.4\textwidth}
    % source code: bivariateSelVSat.R
    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_nullDists.pdf}
    \caption{Conditional distributions of $Y_1$ under
      $M_0:\; Y \sim \cN(0,I_2)$. The realized value $|Y_1|=2.9$ is
      quite large given that $j_1=1$. By
      contrast, $|Y_1|=2.9$ is not especially large once we 
      also condition on $Y_2=2.5$.}
  \label{fig:bv_nullDists}
  \end{subfigure}
  \caption{Contrast between the saturated-model and selected-model
    tests in Example~\ref{ex:bivariate}. The selected-model test is based on  $\L(Y_1 \gv j_1=1)$,  whereas the saturated-model test is based on $\L(Y_1  \gv Y_2, \; j_1=1)$. 
    When $Y=(2.9, 2.5)$, the selected- and saturated-model $p$-values are 0.007 and 0.3, respectively.}
  \label{fig:bv}
\end{figure}

@@

Until this point, we have deferred discussing how to compute the reference distribution for the max-$t$ test, next-entry test, or any of the other tests we have discussed. In each case, the $p$-value at step $k$ comes from a test that rejects when some test statistic $T_k$ is large compared to the law
\begin{equation}\label{eq:genDist}
\L(T_k(Y) \,\mid\, M_{k-1}(Y), U_{k-1}(Y)),
\end{equation}
under the null hypothesis that $F\in M_{k-1}$. Because $U_{k-1}$ is sufficient for $Y$ under the null, this law is completely known. Thus, we have reduced computation of the $p$-value at step $k$ to a sampling problem: if we can sample values of $Y$ from its null distribution given $M_{k-1}(Y), U_{k-1}(Y)$, and compute the test statistic $T_k(Y)$ for each sample, we can numerically approximate the $p$-value to whatever precision we desire.

@@

At each step we use the selective max-$z$ test conditioning on the current null model, where $p_k$ is small when $|z_{E_{k-1}}^*|$ is large compared to the law
\[
\L(|z_{E_{k-1}}^*| \,\mid\, E_{k-1}, Y_{E_{k-1}}).
\]

>>>>

\newcommand{\cM}{\mathcal{M}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\FDR}{\textnormal{FDR}}
\newcommand{\FCR}{\textnormal{FCR}}
\newcommand{\crt}{\phi}
\newcommand{\M}{\mathcal{M}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\Gv}{\;\;\big|\;\;}
\newcommand{\proj}{\cP}
\newcommand{\pow}{\text{Pow}}
\newcommand{\supp}{\text{supp}}
\newcommand{\sF}{\mathscr{F}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\sC}{\mathscr{C}}
\newcommand{\hJ}{\widehat{J}}
\newcommand{\bH}{\mathbf{H}}
\newcommand{\bM}{\mathbf{M}}
\newcommand{\tM}{\widetilde{M}}
\newcommand{\tE}{\widetilde{E}}
\newcommand{\tV}{\widetilde{V}}
\newcommand{\tR}{\widetilde{R}}
\newcommand{\tL}{\widetilde{L}}
\newcommand{\hk}{\hat{k}}
\newcommand{\hr}{\hat{r}}       
\newcommand{\cN}{\mathcal{N}}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}
\newcommand{\Err}{\mathcal{E}}
\newcommand{\RSS}{\text{RSS}}

@@

Following \citet{fithian2014optimal}, we say that for a fixed candidate null model $M$, the random variable $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is stochastically larger than uniform (henceforth super-uniform) under sampling from any $F\in M$, given that $M$ is selected. That is,
\begin{equation*}
\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = M\right) 
\leq \alpha, \quad \forall F\in M, \; \alpha \in [0,1].
\end{equation*}
Once we have constructed selective $p$-values for each fixed $M$, we can use them as building blocks to construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
\[
p_k(y) = p_{k, M_{k-1}(y)}(y),
\]
which is super-uniform on the event $\{F \in M_{k-1}(Y)\}$:
\begin{equation}\label{eq:selectiveGuaranteePk}
\P_F\left(p_k \leq \alpha \mid F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
\end{equation}
One useful strategy for constructing exact selective tests is to condition on the sufficient statistics of the null model $M_{k-1}$. By sufficiency, the law
\[
\cL_F(Y \;\mid\; U_{k-1}, \; M_{k-1}=M)
\]
is the same for every $F\in M$. Thus, we can construct selective tests and $p$-values by comparing the realized value of any test statistic $T_k(Y)$ to its known conditional distribution under the null. It remains only to choose the test statistic and compute its conditional null distribution, which can be challenging. See \citet{fithian2014optimal} for a general treatment. Section~\ref{sec:computation} discusses computational strategies for the tests we propose in this article.

@@

A natural choice of test statistic is the {\em greedy likelihood ratio statistic}
\begin{equation}\label{eq:greedyLRT}
G_k(Y) = \sup_{\theta\in \Theta(E_k)} {\ell(\theta; Y)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell(\theta; Y)},
\end{equation}
which is the generalized likelihood ratio statistic for testing $M(E_{k-1})$ against the ``greedy'' alternative with one more active parameter, $\bigcup_{j \notin E_{k-1}} M(E_{k-1} \cup \{j\})$. The {\em selective greedy likelihood ratio test} rejects for large $G_k$, based on the law
\begin{equation*}
\cL\left(G_k \;\mid\; M_{k-1}, U_{k-1}\right)
\end{equation*}
Because the likelihood in linear regression is a monotone decreasing function of the residual sum of squares, the max-$t$ test is equivalent to the greedy likelihood ratio test. The counterpart of the max-$t$ test in linear regression with known $\sigma^2$ is the max-$z$ test, which differs only in replacing the $t$-statistics $t_{j,E}$ with their corresponding $z$-statistics. The max-$z$ test is also equivalent to the selective greedy likelihood ratio test.

@@

In this setting, $\lambda_k$ is a natural test statistic for model $M_{k-1}$, with larger values suggesting a poorer fit. The {\em selective next-entry test} is the test that rejects for large $\lambda_k$, based on the law
\begin{equation*}
\cL\left(\lambda_k \;\mid\; U_{k-1}, M_{k-1}\right).
\end{equation*}

@@

We now review several proposals for ordered hypotheses testing along with their error guarantees in the traditional setting, where the sequence of null hypotheses is fixed. In the next section we will extend the analysis to random null hypotheses.

@@

A proof of Proposition~\ref{prop:counterexample} is given in the appendix.  The problem is that $k_0$ is no longer a fixed index. Consequently, even though $\cL(p_k \mid H_k \text{ true})=U[0,1]$ for each fixed $k$, the $p$-value \smash{$p_{k_0}$} corresponding to the {\em first} true null hypothesis is stochastically smaller than uniform. In this counterexample, $k_0=2 \iff |Y_3|>z_{\alpha/2}$, giving $\text{FWER}= 1$ conditional on the event $k_0=2$ and leaving no room for error when $k_0\neq 2$.

@@

Similarly, the error control guarantees of \citet{gsell2013sequential} and \citet{li2015accumulation} do not directly apply to case where the null hypotheses are random. However, we recover these guarantees if we have {\em conditional} independence and uniformity of null $p$-values given $k_0$: that is, if for all $k=0,\ldots, d-1$ and $\alpha_{k+1},\ldots,\alpha_{d} \in [0,1]$, we have
\begin{equation}\label{eq:indepCond_random_k0}
  \P_F(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
  \mid p_1, \ldots, p_k) \eqAS \left(\prod_{i=k+1}^d \alpha_i\right) \quad \text{on } \{k_0(Y, F) = k\},
\end{equation}
For the sake of brevity, we will say that $p$-value sequences satisfying~\eqref{eq:indepCond_random_k0} are {\em independent on nulls}.

The following proposition shows that independence on nulls allows us to transfer the error-control guarantees of ForwardStop and accumulation tests to the random-hypothesis case.
\begin{proposition}
  Let $\hk$ be a stopping rule operating on $p$-values $p_{1:d}(Y)$ for nested hypotheses $H_{1:d}(Y)$. For some function $g$, let
  \[
  \Err_g = \E\left[ g\left(\hk(p_{1:d}(Y)), k_0(Y, F)\right)\right],
  \]
  Suppose that \smash{$\hk$} controls $\Err_g$ at level $\alpha$ if $H_{1:d}$ are fixed and the null $p$-values are uniform and independent of each other and the non-nulls. Then, \smash{$\hk$} controls $\Err_g$ at level $\alpha$ whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
\end{proposition}
\begin{proof}
    For nested hypotheses, $k_0$ completely determines the truth or falsehood of $H_k(Y)$ for every $k$. If \smash{$\hk$} controls $\Err_g$ in the 
fixed-hypothesis case, it must in particular control $\Err_g$ conditional on any fixed values of $p_{1:k_0}$, since we could set 
\[
p_{1:k_0} \sim \prod_{k=1}^{k_0} \delta_{a_k}
\]
for any sequence $a_1,\ldots, a_{k_0}$, where $\delta_a$ is a point mass at $a \in [0,1]$.

Thus, \eqref{eq:indepCond_random_k0} implies
  \[
  \E\left[ g\left(\hk(p_1,\ldots,p_d), k_0(Y, F)\right) \;\mid\;
      k_0(Y, F), \; p_{1:k_0}(Y)\right] \leqAS \alpha.
  \]
  Marginalizing over $(p_{1:k_0},k_0)$ gives $\Err_g\leq \alpha$.
\end{proof}

@@

Recall that at step $k$, the max-$t$ test rejects for large \smash{$T_k=|t_{E_{k-1}}^*|$}, comparing its distribution to the conditional law
\begin{equation}\label{eq:maxTnull}
\cL\left(T_k \; \mid \; E_{k-1}, \,X_{E_{k-1}}'Y, \,\|Y\|^2 \right).
\end{equation}
This conditional distribution is the same for all $F\in M(E_{k-1})$, because \smash{$U_{k-1}=(X_{E_{k-1}}'Y,\; \|Y\|^2)$} is the complete sufficient statistic for $M(E_{k-1})$.

If $F\in M(E_{k-1})$, then $p_k$ is uniform and independent of the previous $p$-values $p_{1:(k-1)}$ since:
\begin{enumerate}
\item $p_k$ is uniform conditional on $E_{k-1}$ and $U_{k-1}$ by construction, and
\item $p_1, \ldots, p_{k-1}$ are functions of $E_{k-1}$ and $U_{k-1}$, as we will see shortly.
\end{enumerate}
Informally, the pair $(E_{k-1}, U_{k-1})$ forms a ``wall of separation'' between $p_k$ and $p_{1:(k-1)}$, guaranteeing that $\cL(p_k \mid p_{1:(k-1)}) = U[0,1]$ whenever $H_k$ is true.

@@

Furthermore, for $i<k$, $p_i$ is computed by comparing $T_i=|t_{j_i, E_{i-1}}|$, which is a function of $(E_i, U_i)$, to the reference null distribution $\cL_{H_i}(T_i \mid E_{i-1}, U_{i-1})$, which is a function of $(E_{i-1}, U_{i-1})$. Having verified that under $H_k$, $p_k$ is uniform and independent of $p_{1:{k-1}}$, we can apply this conclusion iteratively to see that all remaining $p$-values are also uniform and independent.

By contrast, the saturated model $p$-values are computed using a reference distribution different from~\eqref{eq:maxTnull}, one that depends on
information not contained in $(E_k,U_k)$. As a result, saturated-model $p$-values are generally not independent on nulls. We discuss the regression setting in more detail in Section \ref{sec:selective-reg}.

@@

\begin{corollary}\label{cor:greedyLikIndep}
  If $U(Y; M)$ is complete and sufficient for model $M$, then the selective greedy likelihood $p$-values computed for the forward stepwise path are independent on nulls.
\end{corollary}
\begin{proof}
  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the sufficient filtration for the forward stepwise path. First, the test statistic $G_k$ is $\sF_k$-measurable because
  \begin{align*}
    G_k(Y) &= \sup_{\theta\in \Theta(E_k)} {\ell(\theta; Y)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell(\theta; Y)}\\
    &= \sup_{\theta\in \Theta(E_k)} {\ell_{E_k}^{Y\mid U}(\theta; U_k)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell_{E_k}^{Y\mid U}(\theta; U_k)}.
  \end{align*}
  Second, the null distribution $\cL\left(G_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $G_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
\end{proof}

@@

  Furthermore, for all $\lambda\in\Lambda, \lambda\geq \gamma_k$, we have \smash{$\tE_\lambda \sub E_k$}. As a result, \smash{${\hat\theta^{\lambda} =\hat\theta^{(E_k,\lambda)}}$} for such $\lambda$, so $\tE_{\lambda}$ depends only on \smash{$\ell_{E_k}^{Y\mid U}(Y;\, U_k)$}, and therefore $\lambda_k$ can be computed from $(E_k, U_k)$.  Second, the null distribution $\cL\left(\lambda_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $\lambda_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
\end{proof}

@@

Let $\eta_k\in \R^n$ denote the vector for which \smash{$\theta_{j_k}^{E_k} = \eta_k'\mu$}, and let \smash{$\proj_{\eta_k}^\perp$} denote the projection operator into the space orthogonal to $\eta_k$. The UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution \smash{$\cL\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)$}, or equivalently
\begin{equation}\label{eq:satModel}
\cL\left( X_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)
\end{equation}

@@

\subsubsection{Testing Against the Next Model}\label{sec:identify}
One option for selected-model inference is to perform a selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$, conditioning on both models. Conditioning on $(M_{k-1},M_k)$ is equivalent to conditioning on $(E_{k-1}, j_k)$. If $\sigma^2$ is known, the test is based on
\begin{equation}\label{eq:selModel_cond}
\cL\left(X_{j_k}'Y \mid E_{k-1}, \; j_k, \; X_{E_{k-1}}'Y\right).
\end{equation}
If $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$. We can construct $p$-values from the UMPU, equal-tailed, or one-sided test based on the one-parameter exponential family~\eqref{eq:selModel_cond}; the only difference is in how much of the overall level $\alpha$ is apportioned to the right or left tail \citep{fithian2014optimal}. If we use the construction in~\eqref{eq:selModel_cond} after a subpath-sufficient path algorithm, the resulting $p$-values are always guaranteed to have independence on nulls. This is because the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, as is the reference null distribution.

@@

To avoid conditioning on the next model $M_k$, we could instead test $M_{k-1}$ against the alternative $M_{\infty}\setminus M_{k-1}$. In that case, we could modify~\eqref{eq:selModel_cond} and base the test on
\begin{equation}\label{eq:selModel_marg}
\cL\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right).
\end{equation}
Again, if $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$. We could also replace $X_{j_k}'Y$ with any other test statistic $T(Y)$. As long as $T(Y)$ is measurable with respect to $(M_k, U_k)$, and the path algorithm is subpath-sufficient, the $p$-values are guaranteed to be independent on nulls. The next-entry test for the lasso and the max-$t$ test in forward-stepwise regression with unknown $\sigma^2$ are both examples of this approach. Note that under resampling from the law in~\eqref{eq:selModel_marg}, the selected variable $j_k$ is random; by contrast, \eqref{eq:selModel_cond} conditions on $j_k$.

@@

\begin{figure}
  \centering
  \begin{subfigure}[t]{.4\textwidth}
    % source code: bivariateSelVSat.R
    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_condSets.pdf}
    \caption{The selected-model test conditions on $j_1=1$ (yellow region), while the saturated-model test also conditions on $Y_2=2.5$ to eliminate the nuisance variable $\mu_2$ (brown line segments).}
    \label{fig:bv_condSets}
  \end{subfigure}
  \hspace{.1\textwidth}
  \begin{subfigure}[t]{.4\textwidth}
    % source code: bivariateSelVSat.R
    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_nullDists.pdf}
    \caption{Conditional distributions of $Y_1$ under
      $M_0:\; Y \sim \cN(0,I_2)$. The realized value $|Y_1|=2.9$ is
      quite large given that $j_1=1$. By
      contrast, $|Y_1|=2.9$ is not especially large once we 
      also condition on $Y_2=2.5$.}
  \label{fig:bv_nullDists}
  \end{subfigure}
  \caption{Contrast between the saturated-model and selected-model
    tests in Example~\ref{ex:bivariate}. The selected-model test is based on  $\cL(Y_1 \gv j_1=1)$,  whereas the saturated-model test is based on $\cL(Y_1  \gv Y_2, \; j_1=1)$. 
    When $Y=(2.9, 2.5)$, the selected- and saturated-model $p$-values are 0.007 and 0.3, respectively.}
  \label{fig:bv}
\end{figure}

@@

Until this point, we have deferred discussing how to compute the reference distribution for the max-$t$ test, next-entry test, or any of the other tests we have discussed. In each case, the $p$-value at step $k$ comes from a test that rejects when some test statistic $T_k$ is large compared to the law
\begin{equation}\label{eq:genDist}
\cL(T_k(Y) \,\mid\, M_{k-1}(Y), U_{k-1}(Y)),
\end{equation}
under the null hypothesis that $F\in M_{k-1}$. Because $U_{k-1}$ is sufficient for $Y$ under the null, this law is completely known. Thus, we have reduced computation of the $p$-value at step $k$ to a sampling problem: if we can sample values of $Y$ from its null distribution given $M_{k-1}(Y), U_{k-1}(Y)$, and compute the test statistic $T_k(Y)$ for each sample, we can numerically approximate the $p$-value to whatever precision we desire.

@@

At each step we use the selective max-$z$ test conditioning on the current null model, where $p_k$ is small when $|z_{E_{k-1}}^*|$ is large compared to the law
\[
\cL(|z_{E_{k-1}}^*| \,\mid\, E_{k-1}, Y_{E_{k-1}}).
\]
